The Calibration Reference Data System

### 2015-10-20 docs/paper/crds.tex
latest, maybe last presubmission version?
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2137 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

We describe a software architecture and implementation for using rules to
determine which calibration files are appropriate for use in calibrating a
given observation. This new system replaces what had been previously used for
the Hubble Space Telescope (HST) calibration pipelines, a system called the
Calibration Database System (CDBS), and will be used for the James Webb Space
Telescope (JWST) calibration pipelines, and is currently being used for HST calibration pipelines. . The approach used for this can be
easily generalized for use in similar applications that need a rules-based
system for selecting the appropriate file for a given case; we give some
examples of such generalizations that will likely be used for JWST. The core
functionality of the Calibration Reference Data System is available under an
Open Source license.
\end{abstract}

@@

When running automatic calibration pipelines one important capability needed is
the ability to identify automatically the appropriate calibration files that
should be used when calibrating the data. The kinds of calibration files are
typically numerous, and there may be many that have to be applied to a specific
dataset. Examples are dark current, bias, flat field, photometric, and
geometric distortions, among many others. Typically calibration files depend on
particular observing modes or parameters, or may be time variable.
Calibration files also differ across instruments both in purpose and specific 
instantiations of a particular reference type.

@@

The system that was used for the Hubble Space Telescope (HST) was called the
Calibration Database System (CDBS,  
\citealt{1994Cox,1997Lubow,1998Cox,1997hstinternal,2004Swam}). It was used successfully for many years (with some design changes documented in the 
cited references). The experience with it revealed that there were some limitations that
proved to constrain its capabilities, and as a result, a newer system that
removed these constraints was sought. We will briefly describe the design of
CDBS at a high level and the constraints that it posed. This provides the
necessary background for discussing the design goals of CRDS.

@@

The most critical aspect of the validation requirements for reference files
concerned the keyword values used to match datasets to the appropriate
references.  Unlike dataset parameter keywords which describe a specific
instrument configuration for one observation, reference file keywords
circumscribe range of parameter values for which the reference applies.  In
CDBS, ranges of parameter values were often represented by intermediate values
which expand into the discrete values used to define specific instrument
configurations.  As part of submitting reference files, the intermediate values
were combinatorially exploded into database rows which represented each
possible application of a reference to some discrete instrument parameter
configuration.  As an example, a dataset APERTURE keyword would describe the
specific aperture an instrument was configured to use for one observation. In
contrast, the same APERTURE keyword specified in a reference file would
describe, indirectly, the combined set of different specific APERTUREs that
reference file was capable of supporting.  When multiple parameters were used
to match datasets to references, the number of distinct rows in the CDBS
database was equal to the product of the number of specific values supported by
a particular reference for each parameter.

@@

\begin{enumerate}
\item Difficulty testing new reference files. The use of a database essentially
limited the system to one set of rules for selecting reference files. Once a
reference file was delivered to the system, effectively the rules for selection
were modified in the operation system immediately. There were occasions when
reference files were deemed to have passed functional testing (i.e., they
performed the correct calibration for the designated cases), but were submitted
with incorrect information about which modes or dates that they should be used
for. As a result, the operational pipeline would only discover that there was a
problem when processing real data, thus resulting in dealing with correcting
mis-processed data (or in some cases, failure to even process). Problems in the
operational pipeline are significant disruptions. In the last few years, a test
version of the database was created so that tests could be performed, but that
introduces synchronization and management issues to make sure what is being
tested corresponds to the previous operational version. Furthermore, if
multiple deliveries are in progress from different instruments, it introduces a
coordination issue to make sure they are either all completed together or in
sequence, which can prove constraining to the needs of each instrument as far
as testing goes.

\item Difficulty in undoing mistakes. A delivery of a file with mistaken
parameters could be difficult to undo. No reference file could be removed from
the database for various reasons. In a number of cases, particularly if the
date that it was targeted for was incorrect, it could seriously corrupt the
selection logic. To take the most common error case: Files often had a
``useafter'' date associated with them that indicated that it should be used for
a date after the specified one *if no other file had a later date that was
still before the date of the observation*. If one submitted a file intended to
be used for data after 2000-01-01 but mistakenly provided
1999-01-01 then two new file submissions are required to correct the problem.
One to resubmit the new file with the correct date; the second to resubmit the file
intended to be used after 1999-01-01 so that its useafter date is one second after
the mistaken submission.

@@

\item Difficulty in supporting remote usage by astronomers. 
Using the bestref facility remotely
requires providing a web service so that remote processing can get the most up
to date recommendations on the best reference file to use. There is one serious
drawback in this though. Frequently the versions of the calibration software in
the operations pipeline have not yet been publicly released, and there are
times when the latest reference files require the latest software. Such a web
service has the potential of recommending reference files that either are
inappropriate for previous versions of the calibration pipeline software, or
simply will not work with previous versions. As a result, this service was not
provided.

\item Difficulty in allowing customized variants of selection rules. Sometimes
observers or teams have specialized calibration files that they wish to
substitute in place of the standard ones. There was no simple way of allowing a
team to run a customized version of CDBS to support this. The entrenched
software requirements were too numerous (when one develops a system to run in
only one environment, dependencies on that environment easily become entrenched
in the system). Even if no customization was desired, it makes running the
calibration pipeline remotely at another institution very difficult since there
is no simple way to get the the most recent recommendations.

\item Difficulty in providing remote pipelines a consistent bestref
environment. There are times observers *do not* want changing references files
if they wish comparison to previously computed results. They want the same
rules applied, even if they aren't the very best version.
As a ``single state'' database expressing only the current best reference
assignment rules, reproducing historical results in CDBS was difficult or
impossible.

\item Difficulty in understanding what the effective rules are. The exact
selection rules are embodied in a database with a history of transactions for
which many supersede previous ones. One really only knows the net effect by
running queries on the database for specific cases, even when something quite
simple could summarize what is desired for the current situation. The net
effect of this is that people do not really understand what the rules are, and
that they are not what was intended. Finding mistakes in this situation can be
quite difficult. And once found, can be difficult to rectify.

\item Difficulty in adding new kinds of selection rules. The database schema
effectively constrains what kinds of rules can be used. More complex rules
either lead to horrendously complex queries, or custom software to generate
custom queries. The use of the custom software/queries ends up making the system
even more opaque.
\end{enumerate}

@@

\begin{enumerate}
\item Databases are effectively one state systems. Yes, one can deal with
multiple states but usually at great complexity.
\item Databases do not make it simple to distribute software since installing,
administering and maintaining a database elsewhere is a major task, and keeping
them synchronized to the master copy is yet a greater complexity
\item Once established, changing the structure of database is difficult,
particularly with a systems used in operations.
\end{enumerate}

Is it in fact necessary to use a database to hold the rules? In reviewing all
the previous and anticipated cases for HST and JWST, it did not appear that a
database was necessary for selection rules, and that a simpler and more
flexible approach could be chosen. Indeed, all that appears to be needed is to
encompass a version of the rules in a simple text file. In doing so there are a
number of immediate benefits:

\begin{enumerate}
\item the rules are explicit
\item multiple versions of these rules may exist without conflict, and used in
different contexts simultaneously without confusion.
\item these rule files may be easily distributed so that remote users can run the
``bestref'' functionality remotely without needing to install or administer
complex software.
\item the rules are easily customized, even by hand if necessary if
users wish to.
\end{enumerate}

\section{High level organization of CRDS}

The software for CRDS can be considered to be in one of three categories:

\begin{enumerate}
\item supporting the bestref functionality
\item supporting submitting new reference files and all the issues that
go with that
\item Supporting format, semantic, and parameter constraint checking for both
references and rules files.  Capability to detect typos, duplicate lines, and
grammatical errors in rules files.   Capability to analyze basic properties
such duplicate or deleted rows,  by pseudo-mode, for supported tables files.
\item providing utilities for making CRDS useful in operations, for instrument
scientists that must generate and monitor reference files, and end users.
\end{enumerate}

Each of these will be described in a separate section. Since most of the design
is centered around the first category, most of the description will be centered
around that.

@@

The system is centered around rules files that describe (for one version) 
how reference files should be selected for a single type of
reference file. In a general sense, the file consists of a header of sorts that
describes information about what kind of reference file that it is applied to,
what the relevant information from the dataset is needed to make that
selection, and other bookkeeping information. Following that header are the
rules. Generally these are nested sets of criteria that are used to
successively narrow down the selection until a single result is determined. It
is important to understand that there is no intrinsic limit to what structures
can be supported. If more complex cases arise, new structures can be added to
the library to support these. The following lists existing and possible kinds
of selections structures. Those that already are supported are marked with an
`*'. The ones that are not supported are intended to illustrate possible
extensions that are not purely hypothetical. The following examples of files
will illustrate specifically these structures.

\begin{enumerate}
\item matching on sets of discrete parameter values. That is, the item associated
with this specific set of parameters is selected if the parameter values are
matched according to one of several matching approaches. An item may be a file
name, or a nested set of rules that will be evaluated in turn.  
There are a number of different parameter match types, literal, or-glob,
pure regex, between, not, relational, etc.  In the case of multiple parameters,
each parameter is matched across all match cases using a process of elimination
(winnowing) which successively narrows the possible match cases.

\item matching on a sequence of dates. This corresponds to the useafter mechanism
in CDBS. Normally these items will be sorted in date/time order, and the
date/time of the observation is used to select the appropriate item in the list
using a binary search.

\item matching software version requirements. This structure indicates selection
based on versions of the software which item should be selected. It may use
specific versions, or relative (e.g., any version before or up to the specified
version, or any version after the specified version).

\item getting bracketing files. This is useful if some sort of interpolation
between reference files is desired, in which case a pair of results is
returned.

\item matching on ranges of non-discrete values. For example, this may select a
file if an observational parameter is in the specified range. Normally, this
would be a one-dimensional selection, though even more complex multidimensional
volumes could be constructed if need be.
\end{enumerate}

@@

All the files that define the rules for how reference files are selected
use the same organization and syntax.
A simple syntax is used that consists of expected variables set to simple, 
possibly nested, datastructures. Each file is expected to have a header 
and selector defined. The header contains information about the intended 
use of the file, history, information to validate the integrity of the file,
and other similar items. Most important, the header identifies the exposure
parameters that are to be used in the selection process. The selector 
contains the specifics on the rules themselves.

@@

Figure~\ref{fig:example1} illustrates use of the first two items. First note the contents
of the header. It shows what previous rmap file it was derived from, that it is defining
the rules for the selection of reference files (since this syntax is also used for
the selection of other kinds of files), what kind of reference file is being selected, information about the telescope and instrument, and what the name of the file is. 
The {\tt parkey} attribute defines the parameters to use for the selection. In the
case of JWST, the parameters map to FITS header keywords, but are referred through
nested object-oriented scheme; the mapping to specific keywords is handled by a special
library interface that is part of the calibration pipeline. In this specific example,
This is the rules file for selecting the JWST MIRI instrument dark files, based on 
the specific detector, read pattern, and subarray configuration, along with the 
observation date.

@@

This section more systematically describes existing selector mechanisms. 
First is is necessary to understand how the supplied parameters (specified
in the header parkey attribute) are used with selectors. If selectors are 
nested, each layer removes one or more parameters from the list leaving 
the rest for lower nestings. For example, if parkey is as specified in
Figure~\ref{fig:example1}, then 4 parameters are available. The Match 
selector is requesting 3 since all of its keys have a tuple of 3 items,
that leaves only the date parameter for subsequent selectors.
In all cases, the selector expects to be provided an argument that is the 
appropriate type (e.g., if time).

@@

\begin{description}
\item[lists of allowable alternatives:] Allowable values for parameteters
may be supplied as an \textquotesingle OR\textquotesingle 
list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1{\ttfamily"}
| {\ttfamily"}filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt \*} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are braketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37\",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). This does not
affect the weight of the match.
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] A shorthand notation that needs better explanation 
than is in the documenation [I couldn't understatnd it as written xxx]
\end{description}

With all these match options it is possible to write multiple items that 
match a given case. In order to allow effective defaults for general cases
while allowing specific cases to override the general case, special rule
for disambiguation in these situations have been developed. The basic idea
is that more specific matches take precedence over more general matches.
For example, if one item lists as its matching value a wildcard, and another
gives a specific value for the same parameter, the specific match takes
precedence. The precedence is handled by determining the generality of the
match. Use of wildcards or regular expressions increase the generality index,
while use of N/A does not. [I found the explanation of Match weighting
unclear, and it suggests the oppposite, a couple examples would help a lot
 xxx].

If after using these precedence rules, there are still multiple matches,
an error results.

@@

\subsubsection{SelectVersion}
This is essentially a special case of relational matches that understands
the semantics of software versions (e.g., that 1.2 > 1.1.10).

@@

The system is based on these files (which we refer to rmap files since they
indicate the mapping to reference files) but also involves a hierarchy of
related files that are called context files, which provide a nested 
organizational structure corresponding to the functions at different levels. Essentially context files are
text files that define a set of related files that are to be used for
that configuration.
Instrument context files collect all the rmap files for that
instrument. This file effectively defines the versions of all the rmap files
for each reference file types that comprise the version of recommendations in
effect for that instrument. In the example of the instrument context file
below, one can see that each distinct rmap file has a unique name. If anything
is changed in an rmap file, it gets a new name

Figure~\ref{fig:example_instrument_context}

@@

This hierarchy is designed to match that for JWST and as such is currently
designed around the current levels of hierarchy. It is a possible generalization
that would allow it to deal with arbitrary levels of heirarchy, perhaps
different in number depending on the branch taken. Achieving that would
take significant modifications but is possible.

\subsection{Ensuring integrity of the configuration}

@@

\subsection{Practicality and Restrictions of this design}

@@

\begin{enumerate}
\item submit each of the reference files (with validation of each file
being part of the submission process). This is expected to generate
unique names for each of the submitted reference files
\item edit the rmap for that reference file type to include the necessary
changes to use these (most likely either a replacement of older versions
or the addition of time-variable ones).
\item submit the updated rmap file (again with implied validation for
correctness)
and again, generating a new, unique rmap file name that represents the version
of the file in the system that has the desired changes.
\item edit the instrument context file to use the new version of the rmap file.
\item submit the updated instrument context file (with validation and 
generating a new, unique name)
\item edit the pipeline context file to use the new version of the
instrument context file.
\item submit the updated pipeline context file (with validation and 
generating a new, unique name)
\end{enumerate}

@@

Since any number of context can exist, one can have several in testing
simultaneously. If that is the case, one still needs to be careful in
determining if different contexts need merging of information (as could happen
if different instruments are submitting new files independently).

@@

\subsection{Streamlined updates}

@@

In either of these cases, it is quite useful to streamline the process to
effectively make it one step. Namely by the fact of submitting the reference
files, automatically updating and submitting the appropriate rmap to use it,
and in turn, making the corresponding updates and submissions to the context
files. As a result of these submissions, the user then is given all the
relevant names for the new reference files, rmap, and context files, most
importantly the pipeline context file. For this to work properly, the user must
indicate which pipeline context they are basing the changes on. Normally this
is the last one,  known as the ``edit context'' which can be modified several
times in advance of adopting the context as operational in the pipeline.

There are several other options as well:

\begin{itemize}
\item operational
\item one of last 10
\item user specified.  
\end{itemize}

@@

\subsection{Reference File submission details}

@@

File submission is limited to authenticated users. The home page for
authentic pages is shown in Figure~\ref{fig:authenticatedhome}.
The fully automatic
standard process is dubbed ``Batch Reference Submission'' 
in reference to submission of a pile of files for one instrument,
multiple types,  and automatic certification and rules updates.

@@

The authenticated user begains a batch submission of files on the batch
submission page (Figure~\ref{fig:batchsubmission}) 
by identifying the local files designated as the 
files to replace existing ones (note, the local file names are not the 
names the files will have when ingested into CRDS). 
When all the updates have been identified, the submit button [or
whatever] is clicked, and CRDS will then validate each of the reference
files. If they all pass validation, the user will be presented the results of
the results that will be produced (e.g., the new filenames for the reference
files, rmaps and context files, and other summary information). An example of a
successful validation is shown in Figure~\ref{fig:goodsubmissionreview}.
At this point the user confirms
the submission after which they will be assigned unique names, added to the
CRDS database of reference files, sent to the STScI archive for archiving, and
placed in the disk-based repository of reference files. When the submission
process completes, the user is presented with a summary page, shown in
Figure~\ref{confirmedsubmission}

>>>>

We describe a software architecture and implementation for using rules to
determine which calibration files are appropriate for calibrating a
given observation. This new system, the Calibration Reference Data System
(CRDS), replaces what had been previously used for
the Hubble Space Telescope (HST) calibration pipelines, 
the Calibration Database System (CDBS). 
CRDS will be used for the James Webb Space
Telescope (JWST) calibration pipelines, and is currently
being used for HST calibration pipelines. CRDS can be
easily generalized for use in similar applications that need a rules-based
system for selecting the appropriate item for a given dataset; we give some
examples of such generalizations that will likely be used for JWST. The core
functionality of the Calibration Reference Data System is available under an
Open Source license.
\end{abstract}

@@

When running automatic calibration pipelines one important capability needed is
the ability to identify automatically the appropriate calibration files that
should be used when calibrating the data. The kinds of calibration files are
typically numerous, and there may be many that have to be applied to a specific
dataset. Examples include dark current, bias, flat field, photometric, and
geometric distortions, among many others. Typically calibration files depend on
particular observing modes or parameters, or may be time variable.
Calibration files also differ across instruments both in purpose and specific 
instantiations of a particular reference type.

@@

The system that was used for the Hubble Space Telescope (HST) was called the
Calibration Database System (CDBS,  
\citealt{1994Cox,1997Lubow,1998Cox,1997hstinternal,2004Swam}). It was used successfully for many years (with some design changes documented in the 
cited references). The experience with it revealed that there were some limitations that
proved to constrain its capabilities, and as a result a newer system that
removed these constraints was sought. We will briefly describe the design of
CDBS at a high level and the constraints that it posed. This provides the
necessary background for discussing the design goals of CRDS.

@@

The most critical aspect of the validation requirements for reference files
concerned the keyword values used to match datasets to the appropriate
references.  Unlike dataset parameter keywords that describe a specific
instrument configuration for one observation, reference file keywords
circumscribe range of parameter values for which the reference file applies.
In CDBS, ranges of parameter values were often represented by special values
that expand into the discrete values used to define specific instrument
configurations.  As part of submitting reference files, these special values
were combinatorially exploded into multiple database rows that represented each
possible combination of parameter values intended to match to the specified
reference file.  As an example, a dataset APERTURE keyword would describe the
specific aperture an instrument was configured to use for one observation. In
contrast, the same APERTURE keyword specified in a reference file would
describe, indirectly, the combined set of different specific APERTUREs that
reference file was intended to be used for.  When multiple parameters were used
to match datasets to references, the number of distinct rows in the CDBS
database was equal to the product of the number of specific values supported by
a particular reference file for each parameter.

@@

\begin{enumerate}
\item \textbf{Difficulty testing new reference files.} 
The use of a database essentially
limited the system to one set of rules for selecting reference files. Once a
reference file was delivered to the system, effectively the rules for selection
were modified in the operation system immediately. There were occasions when
reference files were deemed to have passed functional testing (i.e., they
performed the correct calibration for the designated cases), but were submitted
with incorrect information about which modes or dates that they should be used
for. As a result, the operational pipeline would only discover that there was a
problem when processing real data, thus resulting in having to deal with correcting
mis-processed data (or in some cases, the failure to even process). Problems in the
operational pipeline are significant disruptions requiring significant work
to correct. In the last few years a test
version of the database was created so that tests could be performed, but that
introduces synchronization and management issues to make sure what is being
tested corresponds to the previous operational version. Furthermore, if
multiple deliveries are in progress from different instruments, it introduces a
coordination issue to make sure they are either all completed together or in
sequence, which can prove constraining to the needs of each instrument as far
as testing goes.

\item \textbf{Difficulty in undoing mistakes.}
A delivery of a file with mistaken
parameters could be difficult to undo. No reference file could be removed from
the database for various reasons. In a number of cases, particularly if the
date that it was targeted for was incorrect, it could seriously corrupt the
selection logic. To take the most common error case: Files often had a
``useafter'' date associated with them that indicated that it should be used for
a date after the specified one \textit{if no other file had a later date that was
still before the date of the observation}. If one submitted a file intended to
be used for data after 2000-01-01 but mistakenly provided
1999-01-01 then two new file submissions are required to correct the problem.
One to resubmit the new file with the correct date; the second to resubmit the file
intended to be used after 1999-01-01 so that its useafter date is one second after
the mistaken submission.

@@

\item \textbf{Difficulty in supporting remote usage by astronomers.} 
Using the bestref facility remotely
requires providing a web service so that remote processing can get the most up
to date recommendations on the best reference file to use. There is one serious
drawback in this though. Frequently the versions of the calibration software in
the operations pipeline have not yet been publicly released, and there are
times when the latest reference files require the latest software. Such a web
service has the potential of recommending reference files that either are
inappropriate for previous versions of the calibration pipeline software, or
simply will not work with previous versions. As a result, this service was not
provided.

\item \textbf{Difficulty in allowing customized variants of selection rules.}
Sometimes
observers or teams have specialized calibration files that they wish to
substitute in place of the standard ones. There was no simple way of allowing a
team to run a customized version of CDBS to support this. The entrenched
software requirements were too numerous (when one develops a system to run in
only one environment, dependencies on that environment easily become entrenched
in the system). Even if no customization was desired, it makes running the
calibration pipeline remotely at another institution very difficult since there
is no simple way to get the most recent recommendations.

\item \textbf{Difficulty in providing remote pipelines a consistent bestref
environment.}
There are times observers \textit{do not} want changing references files
if they wish comparisons to previously computed results. They want the same
rules applied, even if they aren't the very best version.
As a ``single state'' database expressing only the current best reference
assignment rules, reproducing historical results in CDBS was difficult or
impossible.

\item \textbf{Difficulty in understanding what the effective rules are.}
The exact
selection rules are embodied in a database with a history of transactions for
which many supersede previous ones. One really only knows the net effect by
running queries on the database for specific cases, even when something quite
simple could summarize what is desired for the current situation. The net
effect of this is that people do not really understand what the rules are, and
that they are not what was intended. Finding mistakes in this situation can be
quite difficult. And once found, they can be difficult to rectify.

\item \textbf{Difficulty in adding new kinds of selection rules.} 
The database schema
effectively constrains what kinds of rules can be used. More complex rules
either lead to horrendously complex queries, or custom software to generate
custom queries. The use of the custom software/queries ends up making the system
even more opaque.
\end{enumerate}

@@

\begin{enumerate}
\item Databases are effectively one state systems. Yes, one can deal with
multiple states but usually with significantly added complexity.
\item Databases do not make it simple to distribute software since installing,
administering and maintaining a database elsewhere is a major task, and keeping
them synchronized to the master copy is yet a greater complexity.
\item Once established, changing the structure of database is difficult,
particularly with a systems used in operations.
\end{enumerate}

Is it in fact necessary to use a database to hold the rules? In reviewing all
the previous and anticipated cases for HST and JWST it did not appear that a
database was necessary for selection rules, and that a simpler and more
flexible approach could be chosen. Indeed, all that appears to be needed is to
encompass a version of the rules in a simple text file. In doing so there are a
number of immediate benefits:

\begin{enumerate}
\item The rules are explicit.
\item Multiple versions of these rules may exist without conflict, and used in
different contexts simultaneously without confusion.
\item These rule files may be easily distributed so that remote users can run the
``bestref'' functionality remotely without needing to install or administer
complex software.
\item The rules are easily customized, even by hand if necessary if
users wish to.
\end{enumerate}

\section{High-Level Organization of CRDS}

The software for the Calibration Reference Data System (CRDS) 
can be considered to be in one of three categories:

\begin{enumerate}
\item Supporting the bestref functionality.
\item Supporting the submittion of new reference files and all the issues that
go with that.
\item Supporting format, semantic, and parameter constraint checking for both
references and rules files.  Having the capability to detect typos, duplicate 
lines, and grammatical errors in rules files. 
Having the capability to analyze basic properties
such duplicate or deleted rows,  by pseudo-mode, for supported tables files.
\item Providing utilities for making CRDS useful in: operations; for instrument
scientists that must generate and monitor reference files; and end users.
\end{enumerate}

Each of these will be described in a separate section. Since most of the design
is centered around the first category, most of the description will be concentrated
around that.

@@

The system is centered around rules files that describe (for one version) 
how reference files should be selected for a single type of
reference file. In a general sense, the file consists of a header of sorts that
describes information about what kind of reference file that it is applied to,
what the relevant information from the dataset is needed to make that
selection, and other bookkeeping information. Following that header are the
rules. Generally these are nested sets of criteria that are used to
successively narrow down the selection until a single result is determined. It
is important to understand that there is no intrinsic limit to what structures
can be supported. If more complex cases arise, new structures can be added to
the library to support these. The following lists existing selection mechanisms. 
Examples of files will illustrate specifically these structures.

\begin{enumerate}
\item \textbf{Matching on sets of discrete parameter values.}
That is, the item associated
with this specific set of parameters is selected if the parameter values are
matched according to one of several matching approaches. An item may be a file
name, or a nested set of rules that will be evaluated in turn.  
There are a number of different parameter match types: exact, enumerated
list, wildcards,
pure regex expression, range specification, exclusion test, comparison, etc.  
In the case of multiple parameters,
each parameter is matched across all match cases using a process of elimination
(winnowing) which successively narrows the possible match cases.

\item \textbf{Matching on a sequence of dates.}
This corresponds to the useafter mechanism
in CDBS. Normally these items will be sorted in date/time order, and the
date/time of the observation is used to select the appropriate item in the list
using a binary search.

\item \textbf{Matching software version requirements.} 
This structure indicates selection
based on versions of the software which item should be selected. It may use
specific versions, or relative (e.g., any version before or up to the specified
version, or any version after the specified version).

\item \textbf{Identifying bracketing files.}
This is useful if some sort of interpolation
between reference files is desired, in which case a pair of results is
returned.

\item \textbf{Matching on ranges of non-discrete values.}
For example, this may select a
file if an observational parameter is in the specified range. Normally, this
would be a one-dimensional selection, though even more complex multidimensional
volumes could be constructed if need be.
\end{enumerate}

@@

All the files that define the rules for how reference files are selected
use the same organization and syntax.
A simple syntax is used that consists of expected variables set to simple, 
possibly nested, data structures. Each file is expected to have a header 
and selector defined. The header contains information about the intended 
use of the file, history, information to validate the integrity of the file,
and other similar items. Most important, the header identifies the exposure
parameters that are to be used in the selection process. The selector 
contains the specifics on the rules themselves.

@@

Figure~\ref{fig:example1} illustrates use of the first two items. First note the contents
of the header. It shows what previous rmap file it was derived from, that it is defining
the rules for the selection of reference files (since this syntax is also used for
the selection of other kinds of files), what kind of reference file is being selected, information about the telescope and instrument, and what the name of the file is. 
The {\tt parkey} attribute defines the parameters to use for the selection. In the
case of JWST, the parameters map to FITS header keywords, but are referred through a
nested object-oriented scheme; the mapping to specific keywords is handled by a special
library interface that is part of the calibration pipeline. In this specific example,
This is the rules file for selecting the JWST MIRI instrument dark files, based on 
the specific detector, read pattern, and subarray configuration, along with the 
observation date.

@@

This section more systematically describes existing selector mechanisms. 
First is is necessary to understand how the supplied parameters (specified
in the header parkey attribute) are used with selectors. If selectors are 
nested, each layer removes one or more parameters from the list leaving 
the rest for lower nestings. For example, if \texttt{parkey} is as specified in
Figure~\ref{fig:example1}, then 4 parameters are available. The Match 
selector is requesting 3 since all of its keys have a tuple of 3 items,
that leaves only the date parameter for subsequent selectors.
In all cases, the selector expects to be provided an argument that is the 
appropriate type (e.g., time for UseAfter).

@@

\begin{description}
\item[lists of allowable alternatives:] Allowable values for parameteters
may be supplied as an \textquotesingle OR\textquotesingle
{\tt }list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1{\ttfamily"}
| {\ttfamily"}filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt \*} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are braketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). 
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] Substitutions can be defined in the header. Essentially
they result in the replacement of a string in a rules file with a different
string. This is useful in the case the string is complex and must be entered
in many different rules (e.g., it is a combination of many different 
allowable values), or it is more mnemonic to those that must maintain
the rules than the actual header values are.
\end{description}

With all these match options it is possible to write multiple items that 
match a given case. In order to allow effective defaults for general cases
while allowing specific cases to override the general case, special rule
for disambiguation in these situations have been developed. The basic idea
is that more specific matches take precedence over more general matches.
For example, if one item lists as its matching value a wildcard, and another
gives a specific value for the same parameter, the specific match takes
precedence. The precedence is handled by determining the specificity of the
match. Use of specific values, wildcards, or regular expressions increase
the specificity index, while use of N/A does not. If multiple rules match,
the one with the highest specificity index is selected

If after using these precedence rules, there are still multiple matches and 
if the USEAFTER parameter is used, the most appropriate USEAFTER entry is
chosesn. If not, an error results.

@@

\subsubsection{SelectVersion}
This is essentially a special case of relational matches that understands
the semantics of software versions (e.g., that 1.2 $>$ 1.1.10).

@@

The system is based on these files (which we refer to rmap files since they
indicate the mapping to reference files) but also involves a hierarchy of
related files that are called context files, which provide a nested 
organizational structure corresponding to the functions at different levels. Essentially context files are
text files that define a set of related files that are to be used for
that configuration.
Instrument context files collect all the rmap files for that
instrument. This file effectively defines the versions of all the rmap files
for each reference file types that comprise the version of recommendations in
effect for that instrument. In the example of the instrument context file
shown in Figure~\ref{fig:example_instrument_context},
one can see that each distinct rmap file has a unique name. If anything
is changed in an rmap file, it gets a new name

@@

This hierarchy is designed to match that for JWST and as such is currently
designed around the current levels of hierarchy. It is a possible generalization
that would allow it to deal with arbitrary levels of heirarchy, perhaps
different in number depending on the branch taken. Achieving that would
take some modification of the bestref client functionality but is possible
without a large effort. Doing the same for the submission machinery would 
be significantly more work.

\subsection{Ensuring Integrity of the Configuration}

@@

\subsection{Practicality and Restrictions of this Design}

@@

\begin{enumerate}
\item Submit each of the reference files (with validation of each file
being part of the submission process). This is expected to generate
unique names for each of the submitted reference files.
\item Edit the rmap for that reference file type to include the necessary
changes to use these (most likely either a replacement of older versions
or the addition of time-variable ones).
\item Submit the updated rmap file (again with implied validation for
correctness)
and again, generating a new, unique rmap file name that represents the version
of the file in the system that has the desired changes.
\item Edit the instrument context file to use the new version of the rmap file.
\item Submit the updated instrument context file (with validation and 
generating a new, unique name).
\item Edit the pipeline context file to use the new version of the
instrument context file.
\item Submit the updated pipeline context file (with validation and 
generating a new, unique name).
\end{enumerate}

@@

Since any number of contexts can exist, one can have several in testing
simultaneously. If that is the case, one still needs to be careful in
determining if different contexts need merging of information (as could happen
if different instruments are submitting new files independently).

@@

\subsection{Streamlined Updates}

@@

In either of these cases, it is quite useful to streamline the process to
effectively make it a one step operation.
Specifically by the fact of submitting the reference
files, in turn automatically updating and submitting
the appropriate rmap to use it,
and subsequently, making the corresponding updates and submissions to the context
files. As a result of these submissions, the user then is given all the
relevant names for the new reference files, rmap, and context files, most
importantly the pipeline context file. For this to work properly, the user must
indicate which pipeline context they are basing the changes on. Normally this
is the last one,  known as the ``edit context'' which can be modified several
times in advance of adopting the context as operational in the pipeline.

There are several other options for the edit context as well:

\begin{itemize}
\item Operational context.
\item One of last 10 contexts.
\item User specified context.  
\end{itemize}

@@

\subsection{Reference File Submission Details}

@@

File submission is limited to authenticated users. The home page for
authentic pages is shown in Figure~\ref{fig:authenticatedhome}.
The fully automatic
standard process is dubbed ``Batch Reference Submission'' 
in reference to submission of a group of reference 
files for one instrument,
multiple types,  and automatic certification and rules updates.

@@

The authenticated user begains a batch submission of files on the batch
submission page (Figure~\ref{fig:batchsubmission}) 
by identifying the local files designated as the 
files to replace existing ones (note, the local file names are not the 
names the files will have when ingested into CRDS). 
When all the updates have been identified, the submit button is clicked,
and CRDS will then validate each of the reference
files. If they all pass validation, the user will be presented the results of
the results that will be produced (e.g., the new filenames for the reference
files, rmaps and context files, and other summary information). An example of a
successful validation is shown in Figure~\ref{fig:goodsubmissionreview}.
At this point the user confirms
the submission after which they will be assigned unique names, added to the
CRDS database of reference files, sent to the STScI archive for archiving, and
placed in the disk-based repository of reference files. When the submission
process completes, the user is presented with a summary page, shown in
Figure~\ref{fig:confirmedsubmission}

If the validation does not pass, 
the results indicate what the problem is so the user may make the
necessary corrections. Figure~\ref{fig:failedsubmission}
shows such a summary of problems.
